---
title: "p8105_hw2_ym2771"
author: "Yuqi Miao"
date: "9/24/2019"
output: github_document
---

# Problem 1

## Read and clean the Mr. Trash Wheel sheet:
```{r message=FALSE}
library(tidyverse)
library(readxl)
```

```{r}
Mr_trash_wheel <- 
    read_excel('data/Trash-Wheel-Collection-Totals-8-6-19.xlsx', sheet = 1, range = "A2:N408") %>% 
    janitor::clean_names() %>%
    drop_na() %>% 
    mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))

    
## omit rows that do not include dumpster-specific data     
  
## round the number of sports balls to the nearest integer and converts the result to an integer variable?    
```

### Data description:

This dataset has `r nrow(Mr_trash_wheel)` observations and `r ncol(Mr_trash_wheel)`varaibles, and giving information on the dumpter number(`r colnames(Mr_trash_wheel)[1]`), date of collection(`r colnames(Mr_trash_wheel)[2:4]`), amount of total litter(`r colnames(Mr_trash_wheel)[5:6]`), litter type(`r colnames(Mr_trash_wheel)[c(7:length(Mr_trash_wheel)-1)]`) and efficiency of generating power(`r colnames(Mr_trash_wheel)[length(Mr_trash_wheel)]`). 

## Read and clean precipitation data for 2017 and 2018

```{r}
precip_2017 <- 
    read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 6, range = "A2:B14") %>% 
    mutate(Year = 2017) %>% 
    select(Year, everything()) %>% 
    drop_na()
    
precip_2018 <- 
    read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, range = "A2:B14") %>% 
    mutate(Year = 2018) %>% 
    select(Year, everything()) %>% 
    drop_na()
```

## Next, combine precipitation datasets and convert month to a character variable

```{r}
precip_data <- 
    rbind(precip_2017,precip_2018) %>% 
    mutate(Month = month.name[Month])
```

This dataset contains `r nrow(precip_data)` observations and `r ncol(precip_data)` variables, giving information on monthly total precipitation (key variable name: `r names(precip_data)[3]`) in 2017-2018.

* The total precipitation in 2018 is `r sum(precip_2018 %>% pull(Total))`

* The median number of sports balls in a dumpster in 2017 is `r median(Mr_trash_wheel %>% pull(sports_balls))`


# Problem 2

## First, clean the data in pols-month.csv.

```{r}
pols_month <- 
    read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>% 
    separate(col = mon, into = c("year","month","day"), sep = "-") %>%
    mutate(year = as.numeric(year),month = as.numeric(month)) %>% 
    arrange(year, month) %>% 
    mutate(month = month.name[month]) %>%
    pivot_longer(cols = c(prez_gop, prez_dem),names_to = "president", names_prefix = "prez_",values_to = "value") %>% 
    filter(value != 0) %>% 
    select(-value,-day)



```
### discussion:

In variable prez_gop, there exists 5 abnormal values: 2, which is not defined in the data description. In this data manipulation, I regard observations with 2 in prez_gop as the president was republican on the associated date since the corresponding value in var prez_dem suggests the president was not democratic.

## Second, clean the data in snp.csv
```{r}
snp <- read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
    separate(col = date, into = c("month","day","year"), sep = "/") %>%
    mutate(year = as.numeric(year),month = as.numeric(month)) %>% 
    arrange(year, month) %>% 
    mutate(month = month.name[month]) %>%
    select(year, month, everything(), -day)


    
```


## Third, tidy the unemployment data

```{r}
unemployment <- 
    read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
    mutate(year = Year) %>%
    pivot_longer(cols = Jan:Dec, names_to = "month", values_to = "unemployment_rate") %>%
    mutate(year = as.numeric(year), month = match(month, month.abb)) %>% 
    arrange(year,month) %>% 
    mutate(month = month.name[month]) %>%
    select(year, month, everything(),-Year)

```

## Finally, merging data

```{r}
FiveThirtyEight <- 
    left_join(pols_month, snp)
FiveThirtyEight <- 
    left_join(FiveThirtyEight, unemployment) 

```

* The first data frame "pols_month" is the basic data frame with `r nrow(pols_month)` rows * `r ncol(pols_month)` columns, the range of year is from `r range(pols_month %>% pull(year))[1]` to `r range(pols_month %>% pull(year))[2]`, and the names of the key variables are `r names(pols_month)[3:9]`

* The second data frame "snp" has the dimension of `r nrow(snp)` rows * `r ncol(snp)` columns, the range of year is from `r range(snp %>% pull(year))[1]` to `r range(snp %>% pull(year))[2]`, and the names of the key variables is `r names(snp)[3]`.

* The third data frame "unemployment" has `r nrow(unemployment)` rows * `r ncol(unemployment)` columns, the range of year is from `r range(unemployment %>% pull(year))[1]` to `r range(unemployment %>% pull(year))[2]`, and the names of the key variables is `r names(unemployment)[3]`.

* By merging snp and unemployment data into pols_month, we have the final dataset, FiveThirtyEight, with `r nrow(FiveThirtyEight)` rows * `r ncol(FiveThirtyEight)` columns, the range of year is from `r range(FiveThirtyEight %>% pull(year))[1]` to `r range(FiveThirtyEight %>% pull(year))[2]`, for the range of the year in different set is different, some values are missing. The name of key variables are the combination of above, including `r names(FiveThirtyEight)[3:ncol(FiveThirtyEight)]` .





# Problem 3

## Load and tidy the data.
```{r}
pop_bb_name <- 
    read_csv("data/Popular_Baby_Names.csv") %>% 
    janitor::clean_names() %>% 
    mutate(ethnicity = recode(ethnicity,`ASIAN AND PACIFIC ISLANDER` = "ASIAN AND PACI", `BLACK NON HISPANIC` = "BLACK NON HISP", `WHITE NON HISPANIC` = "WHITE NON HISP")) %>%
    mutate(childs_first_name = str_to_title(childs_first_name)) %>% 
    distinct()
```


Produce a table showing the rank in popularity of the name “Olivia”
```{r}
Olivia <-
    pop_bb_name %>% 
    filter(gender == "FEMALE" & childs_first_name == "Olivia") %>%
    pivot_wider(names_from = year_of_birth, values_from = rank,id_cols = ethnicity)

knitr::kable(x = Olivia, caption = " Table 1: The rank the name “Olivia” as a female baby name over time" )    
```

## Produce a similar table showing the most popular name among male children over time.

```{r}
most_pop_male_name <- 
  pop_bb_name %>% 
  filter(gender == "MALE") %>%
  group_by(ethnicity, year_of_birth) %>%
  summarise(name = childs_first_name[match(max(count), count)]) %>%
  pivot_wider(names_from = year_of_birth, values_from = name, id_cols = ethnicity)
knitr::kable(x = most_pop_male_name, caption = "Table 2: The most popular male name of male baby name over time" )    
```


## plotting
```{r}
plot_data <- 
    pop_bb_name %>% 
    filter(year_of_birth == "2016" & ethnicity == "WHITE NON HISP" & gender == "MALE")
  

g <- ggplot(data = plot_data, aes(x= rank, y = count, label = childs_first_name)) +
    geom_point(color = ifelse(plot_data$rank > 10, "grey50","red")) 
g
```







